Pragmatics account for apparent limits on theory of mind use in adults

### 2016-01-24 writing/cogsci-submission/cogsci_keysar.tex
Add exp. 2 overinf results and update item-wise table

\usepackage{hyperref}
\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{color}
\usepackage{todonotes}
\usepackage{dsfont}
\usepackage{textcomp}

@@

Humans can accurately and intelligently reason about the mental states of other humans. Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}. Children acquire this ability by at least age six \cite{WimmerPerner83_BeliefsAboutBeliefs, WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
%
While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking. In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably. 
In this paper we argue that the apparent failures this view is based on are in fact successes for sophisticated social reasoning. 
In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners expect these more informative utterances, and interpret the actual utterances accordingly---resulting in ``errors''.

The argument of Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate. The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects.
\ndg{it'd be nice to have a grid illustrating one of the conditions to refer to here...}
The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions. For example, the objects in one trial included a medium-sized measuring cup in one location and a smaller measuring cup in another location. The director gave an instruction like `move the large cup up one square,' referring to the medium-sized one. 
Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge. To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs. For example, imagine a much larger measuring cup were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `the large cup' to mean this occluded cup (which the director couldn't possibly know about). 
Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent. 
%46\% of participants reached for the hidden object at least twice. 
Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale = .35]{images/overall_view.jpg}
\end{center}
\vspace{-.25cm}
\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
\label{fig:interface}
\end{figure*}

@@

We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g. a battery); we used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Each participant received half the items in the experimental condition and half in the baseline condition. The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items. All object sets, object placements, and corresponding instruction sets were the same for all participants.

@@

%The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects. Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.

@@

\begin{figure}[b!]
\begin{center}
\includegraphics[scale=.8]{images/exp1MouseTracking.pdf}
\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
\label{fig:exp1hover}
\end{center}
\end{figure}

@@

\begin{table*}
\begin{center}
\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
 \# incorrect & 1 & 1 & 2 & 3 & 9 & 7 & 1 & 5 \\
 \# correct   & 11&14&10&13&7 &8   &12& 8 \\
\end{tabular}
\caption{Item-wise error analysis for experiment 2}
\label{table:exp2ItemWise}
\end{center}
\end{table*}

\begin{figure}[b!]
\begin{center}
\includegraphics[scale=.9]{images/exp2MouseTracking.pdf}
\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
\label{fig:exp2hover}
\end{center}
\end{figure}
\section{Exp.~2: Unscripted Replication}
\label{sec:Exp2}

@@

\todo[inline]{rdh: should do a basic analysis of over-informativity... Don't forget about the person who always gave directions like ``start at the third square of the first row and move down one square''}

\subsection{Discussion}

>>>>

\usepackage{hyperref}
\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{color}
\usepackage{todonotes}
\usepackage{dsfont}
\usepackage{array}
\usepackage{textcomp}

@@

\begin{figure*}[t!]
\begin{center}
\includegraphics[scale = .35]{images/overall_view.jpg}
\end{center}
\vspace{-.25cm}
\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director. \todo[inline]{TODO: Use images from tape condition, add extra panel to show matcher view.}}
\label{fig:interface}
\end{figure*}

Humans can accurately and intelligently reason about the mental states of other humans. Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}. Children acquire this ability by at least age six \cite{WimmerPerner83_BeliefsAboutBeliefs, WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
%
While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking. In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably.

In this paper we argue that the apparent failures this view is based on are in fact successes for sophisticated social reasoning. 
In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners expect these more informative utterances, and interpret the actual utterances accordingly---resulting in ``errors''.

The argument of Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate. The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects.
\todo[inline]{Refer to updated Fig.~1 with both director screen and matcher screen }
The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions. For example, the objects in one trial included a medium-sized measuring cup in one location and a smaller measuring cup in another location. The director gave an instruction like `move the large cup up one square,' referring to the medium-sized one. 
Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge. To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs. For example, imagine a much larger measuring cup were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `the large cup' to mean this occluded cup (which the director couldn't possibly know about). 
Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent. 
%46\% of participants reached for the hidden object at least twice. 
Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.

@@

@@

We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Each participant received half the items in the experimental condition and half in the baseline condition. The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items. All object sets, object placements, and corresponding instruction sets were the same for all participants.

@@

%The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects. Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
\begin{table*}
\begin{center}
\begin{tabular}{ m{1.5cm} | m{1.5cm} | m{1.5cm} |  r | m{1.6cm} | m{1.4cm} | m{1.3cm} | m{1.3cm} | m{1.5cm} |}
& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\\hline
 \# incorrect & 1 & 1 & 2 & 3 & 9 & 7 & 1 & 5 \\
 \# correct   & 11&14&10&13&7 &8   &12& 8 \\
\end{tabular}
\caption{Item-wise error analysis for experiment 2}
\label{table:exp2ItemWise}
\end{center}
\end{table*}

@@

@@

@@

Next, we test whether these reductions in error rate and hover-time interference are in fact due to different speaker behavior. Two coders were given the list of messages that directors sent to convey instructions for critical items, blind to condition. The coders determined (1) whether or not the label used by the speaker fit the hidden distractor and (2) whether or not the label used by the director was more or less precise than the scripted label used in Exp.~1. Inter-rater reliability was relatively high, with $r=0.82$ for distractor fit and $r=0.69$ for the precision comparison. Binary rater decisions were averaged to form a single distractor fit score for each message, and a single precision score.

We found that the referring expressions speakers used to convey critical instructions only fit the distractor 53\% of the time, compared to the scripted message in Exp.~1, which fit the distractor 100\% of the time. Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.\~1. As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$. Finally, we test whether the speaker's choice of label accounts for variability in error rates across different items. We find that error rates are higher when the speaker's label fits the hidden distractor $b = 0.79, t(5) = 2.8, p = 0.36$ \todo[inline]{rdh: this isn't the right regression... I did it on the aggregated scores, and it should really be done on the raw subject-by-subject data. It's just a bit tricky to join the error data and the message data.}

\begin{figure}[b!]
\begin{center}
\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
\label{fig:exp2hover}
\end{center}
\end{figure}
\section{Exp.~2: Unscripted Replication}
\label{sec:Exp2}

\begin{figure}[b!]
\begin{center}
\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
\label{fig:exp1hover}
\end{center}
\end{figure}
\subsection{Discussion}
